---
title: "Advanced R - Data Structure"
author: "Byron Tang"
date: "May 12, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Vectors

# 1. What are the six types of atomic vector? How does a list differ from an atomic vector?

Logical, Integer, Double, Character, Complex, and Raw

A list could be recursive, which means a list can contain other list, while an atomic vector is always flat. In addition, a list could contain contents in different types, but an atomic vector must have all contents in the same type. Finally, lists are usually used to build up more complexed data structures in R. For example, data frames and linear model objects.

# 2. What makes is.vector() and is.numeric() fundamentally different to is.list() and is.character()?

is.vector() does not test if an object is a vector. Instead it returns TRUE only if the object is a vector with no attributes apart from names. is.vector() would return TRUE even when we pass a list object into it. 

On the other hand, is.list() tests if an object is build on a list structure. is.list() would return TRUE for a list and FALSE for a vector.

To test if an object is an atomic vector, we need to use is.atomic() || is.list(). 

is.numeric() is a general test for the "numberliness" of a vector and returns TRUE for both integer and double vectors. It is not a specific test for double vectors, which are often called numeric. 

is.character() simply test if an object is of type character or not.


# 3. Test your knowledge of vector coercion rules by predicting the output of the following uses of c():

```{r}
c(1, FALSE) # double
c("a", 1) # character
c(list(1), "a") # X: character O: list 
c(TRUE, 1L) # integer
```

# 4. Why do you need to use unlist() to convert a list to an atomic vector? Why doesn't as.vector() work?

```{r}
# as.vector does not convert a list as a list is also a vector. 
a_list <- list(1, 2, 3)
unlist(a_list)
as.vector(a_list)
is.vector(a_list)
```

# 5. Why is 1 == "1" true? Why is -1 < FALSE true? Why is "one" < 2 false?

- 1 is coerced into a character
- FALSE is coerced to 0
- 2 is coerced to "2"

# 6. Why is the default missing value, NA, a logical vector? What's special about logical vectors? (Hint: think about c(FALSE, NA_character_).)

A logical datatype could be coerced to anyother data types, including integer, logical, and character

```{r}
typeof(c(FALSE, NA))
typeof(c(1L, NA))
typeof(c(1, NA))
typeof(c("1", NA))
```

R has no scalar types, so NA is created as a logical vector of lengh one.

## Attributes

Notes

```{r}
y <- 1:10
attr(y, "my_attribute") <- "This is a vector"
# y <- structure(1:10, my_attribute = "This is a vector")

attributes(y) # returns a list of attirbutes
str(attributes(y))
is.vector(y)# Note that is.vector() would return FALSE when there's an attribute other than name is attached
is.vector(1:10)
```

```{r}
names(y)
dim(y)
class(y)
```

# 1. An early draft used this code to illustrate structure():

# structure(1:5, comment = "my attribute")
# [1] 1 2 3 4 5

# But when you print that object you don't see the comment attribute. Why? Is the attribute missing, or is there something else special about it? (Hint: try using help.)

```{r}
# Use comment() to get the comment attribute
comment(structure(1:5, comment = "my attribute"))
```

# 2. What happens to a factor when you modify its levels?

# f1 <- factor(letters)
# levels(f1) <- rev(levels(f1))

When we modify the levels of a factor, we are fundemantally chaning the category of each integer stored in the factor. For example, if 1 was assigned to a, changing the level a to z would make 1 related to z.

```{r}
e1 <- factor(c("a", "b", "c", "a"))
levels(e1)
table(e1)
levels(e1) <- c("b", "c", "a")
table(e1)
levels(e1)
```

# 3. What does this code do? How do f2 and f3 differ from f1?

# f2 <- rev(factor(letters))
# f3 <- factor(letters, levels = rev(letters))

f2: The sequence of the underlining data is changed, but the order of levels does not change, which means the integer is still assigned to the same level.

f3: Assigns the levels starting from z.

In f1, the levels are changed fundemantally, making the integer redirect to different letters. In f2 and f3, neither of the integer were redirected.

## Matrices and arrays

# 1. What does dim() return when applied to a vector?

```{r}
dim(c(1, 2, 3))
str(c(1, 2, 3))
```

# 2. If is.matrix(x) is TRUE, what will is.array(x) return?

```{r}
is.matrix(matrix(1:6, ncol = 2, nrow = 3))
is.array(matrix(1:6, ncol = 2, nrow = 3))
```

# 3. How would you describe the following three objects? What makes them different to 1:5?

```{r}
x1 <- array(1:5, c(1, 1, 5)) # contains an array of 5 matrices of 1*1
x2 <- array(1:5, c(1, 5, 1)) # contains an array of 1 matrices of 1*5
x3 <- array(1:5, c(5, 1, 1)) # contains an array of 1 matrices of 5*1

x1
x2
x3
```

## Data Frames

# 1. What attributes does a data frame possess?

```{r}
attributes(data.frame(a = 1:3, b = c("a", "b", "c")))
```

# 2. What does as.matrix() do when applied to a data frame with columns of different types?

```{r}
# as.matrix() would coerse different data types
as.matrix(data.frame(a = 1:3, b = c("a", "b", "c")))
```

# 3. Can you have a data frame with 0 rows? What about 0 columns?

```{r}
# A data frame with 0 rows
a <- data.frame(a = as.numeric(),
                b = as.character())
a
class(a)
# A data frame with 0 columns
b <- data.frame()
b
class(b)
```

