---
title: "Functional Programming"
author: "Byron Tang"
date: "July 7, 2018"
output: html_document
---

```{r setup, include=FALSE}

```

## Anonymous functions

1. Given a function, like "mean", match.fun() lets you find a function. Given a function, can you find its name? Why doesn't that make sense in R?

```{r, echo=FALSE}
message("If the function is anonymous, we could not find its name in any way. However, if the function is not anonymous, we'll need to give a function by its 'name' in order to find a name, which seems odd. ")
```

2. Use lapply() and an anonymous function to find the coefficient of variation (the standard deviation divided by the mean) for all columns in the mtcars dataset.

```{r}
lapply(mtcars, function(x) sd(x)/mean(x))
```


3. Use integrate() and an anonymous function to find the area under the curve for the following functions. Use Wolfram Alpha to check your answers.

y = x ^ 2 - x, x in [0, 10]
y = sin(x) + cos(x), x in [-??, ??]
y = exp(x) / x, x in [10, 20]

```{r}
integrate(function(x) x ^ 2 - x, 0, 10)
integrate(function(x) sin(x) + cos(x), -pi, pi)
integrate(function(x) exp(x) / x, 10, 20)
```

The answer is double checked with Wolfram Alpha. However, Wolfram Alpha calculated area under x axis as positive, while integrate() count it as negative. Here I use the calculation of integrate as other materials also suggest the area under x asix should be negative. 
Reference: https://amsi.org.au/ESA_Senior_Years/SeniorTopic3/3f/3f_2content_8.html


## Closures

1. Why are functions created by other functions called closures?

```{r, echo=FALSE}
message("Closures get their name because they 'enclose' the environment of the parent function and can access all its variables.")
```

2. What does the following statistical function do? What would be a better name for it? (The existing name is a bit of a hint.)

bc <- function(lambda) {
  if (lambda == 0) {
    function(x) log(x)
  } else {
    function(x) (x ^ lambda - 1) / lambda
  }
}

```{r, echo=FALSE}
message("The function is Box-Cox transformation:
https://en.wikipedia.org/wiki/Power_transform#Box%E2%80%93Cox_transformation

A name that is more informatinve, like BCTransform,  would be better.")
```

3. What does approxfun() do? What does it return?

```{r, echo=FALSE}
message("The function approxfun() returns a function performing (linear or constant) interpolation of the given data points. For a given set of x values, this function will return the corresponding interpolated values. It uses data stored in its environment when it was created, the details of which are subject to change.")
```

4. What does ecdf() do? What does it return?

```{r}
message("Compute an empirical cumulative distribution function, with several methods for plotting, printing and computing with such an 'ecdf' object. The function ecdf() returns an ecdf project, which is also a function.\n")
x <- rnorm(12)
Fn <- ecdf(x)
Fn     # a *function*
message("\nClass of Fn: ")
class(Fn)
```

5. Create a function that creates functions that compute the ith central moment of a numeric vector. You can test it by running the following code:

m1 <- moment(1)
m2 <- moment(2)

x <- runif(100)
stopifnot(all.equal(m1(x), 0))
stopifnot(all.equal(m2(x), var(x) * 99 / 100))

```{r}
# Create the function
moment <- function(m){
  function(x){
    mean((x - mean(x)) ^ m)
  }
}
```

The fucntion successfully passed the tests.

```{r}
# Test
m1 <- moment(1)
m2 <- moment(2)

x <- runif(100)
stopifnot(all.equal(m1(x), 0))
stopifnot(all.equal(m2(x), var(x) * 99 / 100))
```

6. Create a function pick() that takes an index, i, as an argument and returns a function with an argument x that subsets x with i.

lapply(mtcars, pick(5))
- # should do the same as this
lapply(mtcars, function(x) x[[5]])

```{r}
# Create the function
pick <- function(i){
  function(x){
    len <- length(x)
    stopifnot(i <= len)
    x[[i]]
  }
}
```

```{r}
# Test the function
identical(lapply(mtcars, pick(5)), lapply(mtcars, function(x) x[[5]]))
```

## List of functions

1. Implement a summary function that works like base::summary(), but uses a list of functions. Modify the function so it returns a closure, making it possible to use it as a function factory.

```{r}
# Create the funciton (for numeric inputs only)
summary2 <- function(v){

  functions <- list(
    Min = function(x) min(x),
    Qu_1st = function(x) quantile(x, .25),
    Median = function(x) median(x),
    Mean = function(x) mean(x),
    Qu_3rd = function(x) quantile(x, .75),
    Max = function(x) max(x)
    )
 
  if (!is.data.frame(v)){
    df <- data.frame(lapply(functions,
                            function(f) format(round(f(v), 2), nsmall = 2)),
                     stringsAsFactors = FALSE)
    row.names(df) <- deparse(substitute(v))
  } else if(is.data.frame(v)){
    df <- data.frame(lapply(functions,
                            function(f) format(lapply(mtcars, f))),
                     stringsAsFactors = FALSE)
  }
  t(round(data.matrix(df, rownames.force = NA), 2))
}

summary2(mtcars$mpg)
summary2(mtcars)
```

```{r}
# Function Factory
# The behavior of the new function is different from summary2()
# when taking in a data frame as input
summary_fac <- function(f){

  functions <- list(
    Min = function(x) min(x),
    Qu_1st = function(x) quantile(x, .25),
    Median = function(x) median(x),
    Mean = function(x) mean(data.matrix(x)),
    Qu_3rd = function(x) quantile(x, .75),
    Max = function(x) max(x)
    )
  f <- functions[[f]]
}

fun_names <- c("Min", "Max", "Mean")
funcs <- lapply(setNames(fun_names, fun_names), summary_fac)
with(funcs, paste0("Min: ", Min(mtcars$mpg), 
                   "; Max: ", Max(mtcars$mpg), 
                   "; Mean: ", Mean(mtcars$mpg)))
with(funcs, paste0("Min: ", Min(mtcars), 
                   "; Max: ", Max(mtcars), 
                   "; Mean: ", Mean(mtcars)))
```

2. Which of the following commands is equivalent to with(x, f(z))?

x$f(x$z).
f(x$z).
x$f(z).
f(z).
It depends.

```{r, echo=FALSE}
message("x$f(z)")
```

## Case study: numerical integration

1. Instead of creating individual functions (e.g., midpoint(), trapezoid(), simpson(), etc.), we could store them in a list. If we did that, how would that change the code? Can you create the list of functions from a list of coefficients for the Newton-Cotes formulae?

```{r}
# Create function
integrationFuns <- function(f, a, b, n = 10, rule = "midpoint", open = FALSE){
  
  # Create a closure of Newton-cotes
  newton_cotes <- function(coef, open = FALSE) {
    n <- length(coef) + open
    function(f, a, b) {
      pos <- function(i) a + i * (b - a) / n
      points <- pos(seq.int(0, length(coef) - 1))
        
      (b - a) / sum(coef) * sum(f(points) * coef)
    }
  }
  
  # Create a list of functions
  functions <- list(
    midpoint <- function(f, a, b) (b - a) * f((a + b) / 2),
    trapezoid <- function(f, a, b) (b - a) / 2 * (f(a) + f(b)),
    simpson <- function(f, a, b) (b - a) / 6 * (f(a) + 4 * f((a + b) / 2) + f(b)),
    boole <- newton_cotes(coef = c(7, 32, 12, 32, 7), open = open),
    milne <- newton_cotes(coef = c(2, -1, 2), open = open)
  )
  names(functions) <- c("midpoint", "trapezoid", "simpson", "boole", "milne")
  
  # Calculate composite integration
  points <- seq(a, b, length = n + 1)
  area <- 0
  for (i in seq_len(n)) {
    area <- area + functions[[rule]](f, points[i], points[i + 1])
  }
  area
}

# Test function
message("Default to midpoint: ", integrationFuns(sin, 0, pi))
sapply(c("trapezoid", "simpson", "boole"), 
       function(x) integrationFuns(sin, 0, pi, rule = x))
message("milne: ", 
        integrationFuns(sin, 0, pi, rule = "milne", open = TRUE))
```

Reference: https://en.wikipedia.org/wiki/Newton%E2%80%93Cotes_formulas

2. The trade-off between integration rules is that more complex rules are slower to compute, but need fewer pieces. For sin() in the range [0, ??], determine the number of pieces needed so that each rule will be equally accurate. Illustrate your results with a graph. How do they change for different functions? sin(1 / x^2) is particularly challenging.

```{r}
# Create the function
integrationFuns2 <- function(f, a, b, n = 100, accu = 5, increment = 1) {
  check = FALSE
  
  # Use the output of integrate as the benchmark of accuracy
  trueValue <- integrate(f = f, lower = a, upper = b)
  
  while (check == FALSE){
    allValue <- sapply(c("midpoint", "trapezoid", "simpson", "boole", "milne"),
       function(x) integrationFuns(f = f, a = a, b = b, n = n, rule = x))
    
    # Caculate accuracy to the nth decimal point (default to 5)
    compValue <- round(allValue - trueValue$value, accu)
    
    check <- all(compValue[1] == compValue)
    n = n + increment
  }

  list("TrueValue" = trueValue, 
       "FinalValues" = format(round(allValue, accu), nsmall = accu), 
       "Accuracy at decimal point" = accu, 
       "Accurate" = check, 
       "Pieces" = n)

}

# Test the functions 
integrationFuns2(sin, 0, pi)
```

```{r}
# Test sin(1 / x^2)
testF <- function(x) sin(1/x^2)
integrationFuns2(testF, 1, pi, n = 52500, increment = 10)
```

Illustration by a graph is skipped at this moment.