---
title: "Environments"
author: "Byron Tang"
date: "June 2, 2018"
output: html_document
---

```{r setup, include=FALSE}
library(pryr)
```

## Environment basics

1. List three ways in which an environment differs from a list.

```{r, echo=FALSE}
message("1. Every name in an environment is unique.
2. The names in an environment are not ordered (i.e., it does not make sense to ask what the first element of an environment is).
3. An environment has a parent.
4. Environments have reference semantics.")
```

2. If you don't supply an explicit environment, where do ls() and rm() look? Where does <- make bindings?

```{r, echo=FALSE}
message("ls() and rm() default to the current environment if an environment is not specified.

The operators <- assigns into the environment in which they are evaluated.")
```

3. Using parent.env() and a loop (or a recursive function), verify that the ancestors of globalenv() include baseenv() and emptyenv(). Use the same basic idea to implement your own version of search().

## Recursing over environemnts

```{r}
search()
rec_env <- function(env){
  if (is.null(attr(env, "name"))) {
    print(env)
  } else {
    print(attr(env, "name"))
  }
  if (identical(env, baseenv())) 
    return(emptyenv())
  return(rec_env(parent.env(env = env)))
}
# print(environment())
rec_env(environment())
```

## Function environments

```{r}

```

## Binding names to values

## Explicit environments
