---
title: "Functionals"
author: "Byron Tang"
date: "July 16, 2018"
output: html_document
---

```{r setup, include=FALSE}

```

## My first functional: lapply()
## For loop funcitonals

1. Why are the following two invocations of lapply() equivalent?

trims <- c(0, 0.1, 0.2, 0.5)
x <- rcauchy(100)

lapply(trims, function(trim) mean(x, trim = trim))
lapply(trims, mean, x = x)

```{r, echo=FALSE}
message("In both invocations, the trim argument in mean is supplied by each element in trims while the x argument is constantly the x vector. Therefore, the outputs are the same.")
```

2. The function below scales a vector so it falls in the range [0, 1]. How would you apply it to every column of a data frame? How would you apply it to every numeric column in a data frame?

scale01 <- function(x) {
  rng <- range(x, na.rm = TRUE)
  (x - rng[1]) / (rng[2] - rng[1])
}

```{r}
scale01 <- function(x) {
  rng <- range(x, na.rm = TRUE)
  (x - rng[1]) / (rng[2] - rng[1])
}

# Apply the function to every column of a data frame

# Apply to every numeric column in a data frame
```

3. Use both for loops and lapply() to fit linear models to the mtcars using the formulas stored in this list:

formulas <- list(
  mpg ~ disp,
  mpg ~ I(1 / disp),
  mpg ~ disp + wt,
  mpg ~ I(1 / disp) + wt
)

4. Fit the model mpg ~ disp to each of the bootstrap replicates of mtcars in the list below by using a for loop and lapply(). Can you do it without an anonymous function?

bootstraps <- lapply(1:10, function(i) {
  rows <- sample(1:nrow(mtcars), rep = TRUE)
  mtcars[rows, ]
})

5. For each model in the previous two exercises, extract R2 using the function below.

rsq <- function(mod) summary(mod)$r.squared

```{r}
```

## Data structure functionals

```{r}

```

## FUnctional programming

## Mathematical functionals

## Loops that shouldn't be converted to functions

## A family of functions