---
title: "Functions"
author: "Byron Tang"
date: "May 19, 2018"
output: html_document
---

```{r setup}
library(pryr)
```

## Function components

1. What function allows you to tell if an object is a function? What function allows you to tell if a function is a primitive function?

```{r}
is.function(sum)
is.primitive(sum)
is.primitive(summary)
```

2. This code makes a list of all functions in the base package.

objs <- mget(ls("package:base"), inherits = TRUE)
funs <- Filter(is.function, objs)

Use it to answer the following questions:

  a. Which base function has the most arguments?

  b. How many base functions have no arguments? What's special about those functions?

  c. How could you adapt the code to find all primitive functions?

```{r}
objs <- mget(ls("package:base"), inherits = TRUE)
funs <- Filter(is.function, objs) # 1207 functions

# a
funs_arg <- sapply(lapply(names(funs), formals), length)
(most_arg <- names(funs[funs_arg == max(funs_arg)]))
length(formals(most_arg))

# b
no_arg <- funs[funs_arg == 0]
length(no_arg)
no_arg_primitive <- sapply(no_arg, is.primitive)
sum(no_arg_primitive)
## Most of the functions wihtout arguments are primitive

# c
funs_primitive <- Filter(is.primitive, objs) # 183 primitive functions
```

3. What are the three important components of a function?

```{r}
# 1. the body(), the code inside the function.
# 2. the formals(), the list of arguments which controls how you can call the function.
# 3. the environment(), the "map" of the location of the function's variables.
```

4. When does printing a function not show what environment it was created in?

```{r}
# If the environment isn't displayed, it means that the function was created in the global environment.
```

## Lexical Scoping

1. What does the following code return? Why? What does each of the three c's mean?

c <- 10
c(c = c)

```{r}
# It returns a number 10 with a name c. 
# The funciton looks up to find out an object c and get the number that c represents

# The c in c <- 10 represents a number 10
# In c(c = c):
  # First c is a function
  # Second c is name
  # Third c is the c in c <- 10 that represents a number 10

c <- 10
c(c = c)
```

2. What are the four principles that govern how R looks for values?

```{r}
# name masking
# functions vs. variables
# a fresh start
# dynamic lookup
```


3. What does the following function return? Make a prediction before running the code yourself.

f <- function(x) {
  f <- function(x) {
    f <- function(x) {
      x ^ 2
    }
    f(x) + 1
  }
  f(x) * 2
}
f(10)

```{r}
f <- function(x) {
  f <- function(x) {
    f <- function(x) {
      x ^ 2
    }
    f(x) + 1
  }
  f(x) * 2
}
f(10)
```

## Every Opration is a function call

```{r}
x <- NULL
if (!is.null(x) && x > 0) {
  print("test")
}
```

```{r}
add <- function(x) {
  force(x)
  function(y) x + y
}
adders2 <- lapply(1:10, add)
adders2
```



## Function arguments

1. Clarify the following list of odd function calls:

x <- sample(replace = TRUE, 20, x = c(1:10, NA))
y <- runif(min = 0, max = 1, 20)
cor(m = "k", y = y, u = "p", x = x)

```{r}
x <- sample(c(1:10, NA), 20, replace = TRUE)
y <- runif(20, 0, 1)
cor(x, y, use = "pairwise.complete.obs", method = "kendall")
```


2. What does this function return? Why? Which principle does it illustrate?

f1 <- function(x = {y <- 1; 2}, y = 0) {
  x + y
}
f1()

```{r}
# f1 returns 3 because of lazy evaluation:
  # x in x + y is executed first. 
  # When getting x, y is also assigned as 1, 
  # making it unneccessary to execute y = 0
  # Therefore, the output is 2 + 1 = 3
# The code is slightly revised to grab the value of x and y

f1 <- function(x = {y <- 1; 2}, y = 0) {
  print(x) + print(y)
}
f1()
```

```{r}
# To vefity the thought process, change the order of x and y and observe
# The result shows that when y is executed first,
# y = 0 is captured and y <- 0 is ignored.

f2 <- function(x = {y <- 1; 2}, y = 0) {
  print(y) + print(2)
}
f2()
```

3. What does this function return? Why? Which principle does it illustrate?

f2 <- function(x = z) {
  z <- 100
  x
}
f2()

```{r}
# The function returns 100 
# Default arguments are evaluated inside the function;
# The expression depends on the current environment.
  # Even when there is another global variable z exists,
  # x still captures the z inside the function.

z <- 50
f2 <- function(x = z) {
  z <- 100
  x
}
f2()
```

## Sepcial calls

1. Create a list of all the replacement functions found in the base package. Which ones are primitive functions?

```{r}
# objs <- mget(ls("package:base"), inherits = TRUE)
is_replacement <- grepl("<-", names(objs))
sum(is_replacement) # 56 replacement functions
funs_replacement <- objs[is_replacement]
funs_replacement[sapply(funs_replacement, is.primitive)]
```

2. What are valid names for user-created infix functions?

```{r}
# All user-created infix functions must start and end with %
# Ex: %custom%
```

3. Create an infix xor() operator.

```{r}
# Create the function
# Borrow the code from the source code of xor
`%x||%` <- function(x, y) {
  (x | y) & !(x & y)
}

# Test
TRUE %x||% TRUE == xor(TRUE, TRUE)
TRUE %x||% FALSE == xor(TRUE, FALSE)
FALSE %x||% FALSE == xor(FALSE, FALSE)
```

4. Create infix versions of the set functions intersect(), union(), and setdiff().

```{r}
# Create functions
# The body of the following infix functions are the same with that of the orginal functions
`%i%` <- function(x, y){
  y <- as.vector(y)
  unique(y[match(as.vector(x), y, 0L)])
}

`%u%` <- function(x, y){
  unique(c(as.vector(x), as.vector(y)))
}

`%s%` <- function(x, y){
  x <- as.vector(x)
  y <- as.vector(y)
  unique(if (length(x) || length(y)) 
      x[match(x, y, 0L) == 0L]
  else x)
}

# Test
x <- c(1:5)
y <- c(3:7)

x %i% y
x %u% y
x %s% y
```

5. Create a replacement function that modifies a random location in a vector.

```{r}
# Create the function
`r_replacement<-` <- function(x, value) {
  x[sample(c(1:length(x)), 1)] <- value
  x
}

# Run the test 5 times
for (i in 1:5) {
  x <- 1:10
  r_replacement(x) <- 5L
  print(x)
}
```


## Return values

Note: on.exist()

The code in on.exit() is run regardless of how the function exits, whether with an explicit (early) return, an error, or simply reaching the end of the function body.

```{r}
in_dir <- function(dir, code) {
  # set the directory to a new location 
  #capturing the current location from the output of setwd()
  old <- setwd(dir) 
  # use on.exit() to ensure that the working directory is returned 
  # to the previous value regardless of how the function exits
  on.exit(setwd(old))

  force(code)
}
getwd()
in_dir("~", getwd()) # The environment in the function is changed
getwd()
```

```{r}
in_dir_test <- function(dir, code) {
  #old <- setwd(dir)
  #on.exit(setwd(old))

  force(code)
}
getwd()
in_dir_test("~", getwd()) # The environment in the function is setwd(dir)
getwd()
```

1. How does the chdir parameter of source() compare to in_dir()? Why might you prefer one approach to the other?

```{r}
# chdir: 
  # logical; if TRUE and file is a pathname, 
  # the R working directory is temporarily changed to the directory containing file for evaluating.

# 1) Both methods change the directory in the function 
    # but restore it after when function is done execution.
# 2) I would prefer use chdir as it is a simple argument and saves the work on creating a new function.
```

2. What function undoes the action of library()? How do you save and restore the values of options() and par()?

```{r}
# 1) detach:
    # Detach a database, i.e., remove it from the search() path of available R objects. 
    # Usually this is either a data.frame which has been attached or 
    # a package which was attached by library.
```

options()
Reference: https://stackoverflow.com/questions/15946953/how-to-set-r-to-default-options
  
```{r}
# Save
default_options = options()
save(default_options, file = "default_options.rda")

# Restore
load("default_options.rda")
options(default_options)
```

par()
Reference: RDocumentation

```{r}
# Save
op <- par(mfrow = c(2, 2), # 2 x 2 pictures on one plot
          pty = "s")       # square plotting region,
                           # independent of device size

# Restore
## At end of plotting, reset to previous settings:
par(op)
```

3. Write a function that opens a graphics device, runs the supplied code, and closes the graphics device (always, regardless of whether or not the plotting code worked).

```{r}
# Sample codes from
# https://stat.ethz.ch/R-manual/R-devel/library/grDevices/html/dev.html
# x11()
# plot(1:10)
# x11()
# plot(rnorm(10))
# dev.set(dev.prev())
# abline(0, 1) # through the 1:10 points
# dev.set(dev.next())
# abline(h = 0, col = "gray") # for the residual plot
# dev.set(dev.prev())
# dev.off(); dev.off() #- close the two X devices
```

```{r}
graphIt <- function(code){
  x11()
  force(code)
  ans <- readline(prompt="Press anykey to turn off graph.")
  dev.off()
}

graphIt(plot(1:10))
```

4. We can use on.exit() to implement a simple version of capture.output().

capture.output2 <- function(code) {
  temp <- tempfile()
  on.exit(file.remove(temp), add = TRUE)

  sink(temp)
  on.exit(sink(), add = TRUE)

  force(code)
  readLines(temp)
}

capture.output2(cat("a", "b", "c", sep = "\n"))
## [1] "a" "b" "c"

Compare capture.output() to capture.output2(). How do the functions differ? What features have I removed to make the key ideas easier to see? How have I rewritten the key ideas to be easier to understand?

```{r}
# The input of capture.outpu2() is much more limited. For example:
# capture.output2 only accepts does not accept a file link nor a list object

# The features removed include:
  # write output into the file
  # append output into a file
  # capture output from a list

# The simplified codes in capture.output2 show how capture.output opens
# a file in the function and then close it at the end of function execution by
# replacing a file link as a temp file and removing the temp file
# instead of closing the file passed into the function.

```


```{r}
capture.output2 <- function(code) {
  temp <- tempfile()
  on.exit(file.remove(temp), add = TRUE)

  sink(temp)
  on.exit(sink(), add = TRUE)

  force(code)
  readLines(temp)
}
```

```{r}
capture.output(cat("a", "b", "c", sep = "\n"))
capture.output2(cat("a", "b", "c", sep = "\n"))
```


```{r}
capture.output
```

Reference to more input/output functions: https://www.statmethods.net/interface/io.html
